---
title: "Data Visualization Association"
author: "Md Easin Hasan"
date: "4/4/2021"
output: word_document
---

```{r}
setwd("E:/Data Visualization")
data<-read.csv(file="owid-covid-data_new.csv",header=TRUE)
dim(data)
data1 <- data[data$iso_code == "USA", ] 
dat2 <-data1[,c(1,6,8,11,18,20)]
colnames(dat2)

```
```{r}
miss.info<- function(dat, filename=NULL){  
  
  vnames <- colnames(dat); vnames 
  
  n <- nrow(dat)  
  
  out <- NULL  
  
  for (j in 1: ncol(dat)){  
    
    vname <- colnames(dat)[j]  
    
    x <- as.vector(dat[,j]) 
    
    n1 <- sum(is.na(x), na.rm=T)  
    
    n2 <- sum(x=="NA", na.rm=T)  
    
    n3 <- sum(x=="", na.rm=T)  
    
    nmiss <- n1 + n2 + n3  
    
    ncomplete <- n-nmiss 
    
    out <- rbind(out, c(col.number=j, vname=vname, mode=mode(x), n.levels=length(unique(x)), ncomplete=ncomplete, miss.perc=nmiss/n)) } 
  
  out <- as.data.frame(out)  
  
  row.names(out) <- NULL  
  
  return(out) 
  
}  

miss.info(dat2) 
```

```{r}
library(mice)
fit.mice.dat_f <- mice(dat2, m = 1, maxit = 50, method = 'pmm', seed = 500)
dat_f.imputed <- complete(fit.mice.dat_f,1)
datf <- as.data.frame(dat_f.imputed)
```

```{r}
library("corrplot")
dat3 <- datf[,-c(1)]
M <- cor(dat3)
corrplot(M, method="number")

```

From this correlation plot among the selected variables, we can see that total_deaths and total+cases_per_million, and icu_patients and hosp_patients in USA are very strongly correlated.

# Data Association Imporvements:

```{r}
library(ggcorrplot)
#dat4 <- data.frame(dat3)
corr <- round(cor(dat3), 1)
ggcorrplot(corr)
```
Here the red color is indicating the variables are very strongly correlated and blue color is indicating very weekly correlated. 

```{r}
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)

```
From this correlation plot, we can very easily visualize that total_deaths and total+cases_per_million, and icu_patients and hosp_patients in USA are very strongly correlated. Hence, when the hospitalized patients were high the icu_patients also high in USA during the covid-19 pandemic.
